### 2022-09-02 main/chapter_searches.tex
begin frequentist section intro, add some references

The Large Hadron Collider collides $40$~million proton bunches per second,
but \atlas' trigger system accepts only around $1000$ of
those per second~\cite{atlas2020trigger}.
This means that we have a scenario with a large fixed number $N$ of trials
(bunch collisions)
and a small probability of binary acceptance (triggering),
due to a notable collision event.
The summed total number $n$ of accepted events is therefore described by the
Binomial distribution for $n$;
\begin{equation}
\mathrm{Binomial}(n\mid N, p) = p^n (1 - p)^{N - n} \times \binom{N}{n}
,
\end{equation}
which is a product of the probabilities (produce rule) in the sequence of $N$
accept/reject decisions, summed (sum rule) over all permutations of that sequence
that could have led to the total $n$.
With our tiny trigger rate, we are interested in the limit of small
$p$ and large $N$, which result in a mean acceptance number of $\lambda = Np$.
Using the exponential identity of Equation~\ref{eqn:searches_exponential}, one
can show that this limit tends towards the Poisson
distribution~\cite{jaynes2003probability};
\begin{equation}
\label{eqn:searches_poisson}
\mathrm{Poisson}(n\mid \lambda) = \frac{\lambda^n}{n!}\exp(-\lambda)
,
\end{equation}
which is used ubiquitously in the analysis of our histograms.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.95\textwidth]{figures/seacrhes_fisher_likelihood_binomial.png}
\caption[
Ratio thresholds of a Binomial likelihood function, reproduced from Fisher
]{%
Ratio thresholds of a Binomial likelihood function, reproduced from
Fisher~\cite{fisher1956statistical}.
Levels are drawn at $1/2$, $1/5$, and $1/15$ of its maximum, as a practical way
to communicate its profile.
}
\label{fig:searches_fisher_likelihood}
\end{figure}

@@

>>>>

The Large Hadron Collider collides $40$~million proton bunches per second,
but \atlas' trigger system accepts only around $1000$ of
those per second~\cite{atlas2020trigger}.
This means that we have a scenario with a large fixed number $N$ of trials
(bunch collisions)
and a small probability of binary acceptance (triggering),
due to a notable collision event.
The summed total number $n$ of accepted events is therefore described by the
binomial distribution for $n$;
\begin{equation}
\label{eqn:searches_binomial}
\mathrm{binomial}(n\mid N, p) = p^n (1 - p)^{N - n} \times \binom{N}{n}
,
\end{equation}
which is a product of the probabilities (produce rule) in the sequence of $N$
accept/reject decisions, summed (sum rule) over all permutations of that sequence
that could have led to the total $n$.
With our tiny trigger rate, we are interested in the limit of small
$p$ and large $N$, which result in a mean acceptance number of $\lambda = Np$.
Using the exponential identity of Equation~\ref{eqn:searches_exponential}, one
can show that this limit tends towards the Poisson
distribution~\cite{jaynes2003probability};
\begin{equation}
\label{eqn:searches_poisson}
\mathrm{Poisson}(n\mid \lambda) = \frac{\lambda^n}{n!}\exp(-\lambda)
,
\end{equation}
which is used ubiquitously in the analysis of our histograms.

@@

Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
proportion of stuff within a larger whole, which when formalized as a ratio
of measures and has broad applications.
Frequentist theory confines itself to the special case where those measures
are numbers of discrete randomly occuring events, such that a frequentist
probability is a ratio of frequencies:
\begin{equation}
\label{eqn:searches_prob_frequency_ratio}
\freq{b}{a} = \lim_{N \to \infty}\frac{n}{N},
\end{equation}
where $n$ is the number observed to satisfy conditions $a$ and $b$, and
$N$ have satisfied $b$.
From Equation~\ref{eqn:searches_binomial}, this converges to the general
probability in its infinite limit.

Venn~\cite{venn1866logic},
Boole~\cite{boole1854investigation},
Kolmogorov~\cite{
kolomogoroff1933de,
kolomogoroff1950translated
},

\begin{figure}[tp]
\centering
\includegraphics[width=0.95\textwidth]{figures/seacrhes_fisher_likelihood_binomial.png}
\caption[
Ratio thresholds of a Binomial likelihood function, reproduced from Fisher
]{%
Ratio thresholds of a Binomial likelihood function, reproduced from
Fisher~\cite{fisher1956statistical}.
Levels are drawn at $1/2$, $1/5$, and $1/15$ of its maximum.
Their extents could be reported in a table as a practical communication of
the likelihood function's profile.
}
\label{fig:searches_fisher_likelihood}
\end{figure}

@@

\TODO{
on ``likelihood'' models ---
we are instructed to pretend that these are likelihoods on auxiliary data
TODO cite.
One is free to engage in such doublethink, just as one is free to proclaim that
$2 + 2 = 5$.
The fact remains that many constraints do not correspond to predictions of
auxiliary data: counterexamples include
scale variations,
theoretical cross-section uncertainties,
non-closure uncertainties (which cover differences between estimates).
simulated sample camparisons,
statistical errors on weighted simulations,
extrapolation uncertainties,
and
all uncertainties in supersymmetric signal yields
(for which we have yet no data).
}
